Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-02-02 unstructured.tex
2013.10.29 (Alex) Structured section underwent several reworks/rephrasings. Text reviewed. Minor changes to unstructured section.

While surveying efforts to overcome the mismatch problem
in the area of unstructured \p\ systems,
we came to identify four key methodologies 
utilized by the discussed approaches; they are:
\begin{enumerate}[\itshape i\upshape)]
  \item topology adaptation
  \item forwarding optimization
  \item caching and replication, and
  \item landmarking
\end{enumerate}
%
A key characteristic shared by many of the presented techniques 
is that they do not strictly adhere to a single methodology.
As discussed approaches attempt to address the inherently challenging problem 
of topology mismatch, they resort to heuristics and so, 
they do not often furnish ``pure'' solutions (i.e., solutions that exclusively belong to 
one of the $4$ aforementioned methodologies). 
%%AD rephrased as above..
% An important aspect of many implementations, is that since the above
% methodologies use heuristics to solve topology mismatch, an
% intrinsically difficult problem, no single approach can offer a robust solution.
% Thus, many of the algorithms we surveyed are actually combinations of multiple
% such methodologies, in an attempt to refine the results of using just a single one.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For instance, \emph{mOverlay} combines topology adaptation with landmark binning 
while \emph{Gia} weaves together $3$ methodologies:  topology adaptation, 
forwarding optimization and caching. 
This combination inevitably leads to feature/performance trade-offs: for example, 
the effectiveness of a caching/replication component can be undermined
by a continuously adapting overlay that removes important links between peers.
If a technique is to be evaluated for a specific application domain,
its overall approach as well as its underlying used methodologies have to be considered.
%%AD rephrased as above... did not like much this "advised" - is not wrong BUT sounds funny
% For this reason, the reader is advised to critically approach either the overall
% algorithms or the separate methodologies when evaluating them for different
% application domains.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The cluster-based approaches, on the other hand, 
link physically-close nodes to each other. 
\emph{T2MC}, for example, uses trace-route logs and 
\emph{DDNO} exploits domain names to cluster nodes in proximity at peer join.
Further enhancements may include dynamic local restructuring of the overlay
graph through neighbor exchange like in \emph{PROP} or cycle-cut like in
\emph{DCMP} to achieve continuous adaptation throughout a peer's life-cycle.
Unfortunately, commonly used methods for proximity detection across Internet do
not always return
reliable results and therefore, mapping accuracy is not guaranteed. Specifically
for traceroute, its overhead is not negligible and routers or firewalls in a
network may have already been configured to disable traceroute response from
the start. The most problematic aspect of clustering, though, is
its nature per se. 
Limited connectivity among the various local domains can
significantly shrink the search scope, negatively affecting the query response
time that the \p\ user experiences. Among others, \emph{DDNO} tries to 
balance the efficiency of the clustering approach 
with enhanced node connectivity by forcing half of each node's connections to be
with other, randomly selected, nodes.

As a methodology, topology adaptation, ultimately aims to reduce the 
average path traversing cost from one node to another.
In doing so, the methodology re-arranges the
overlay network so that it becomes a better fit 
for the underlying \emph{IP} network
as Figure~\ref{figure:topology-adaptation} depicts.
%%AD u always need a reference of a Figure in the text!!!
% as Figure~\ref{figure:topology-adaptation} depicts.
Despite the fact that the above is a plausible proposition,
it is not always advantageous though. For example, a topology adaptive
algorithm can exchange a slow edge in the overlay with a faster one
thus reducing the average latency of the network communication.
The pitfall here is that this new virtual link
can traverse a fast \emph{AS}--to--\emph{AS} link meaning that even though
message round-trip-time is reduced, it has actually additional cost in terms of
inter-ISP communication accounting and management.
%%
\begin{figure}[ht]
\centering
\subfigure[Inefficient overlay topology averaging $\simeq 16$ delay units.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
  \label{figure:topology-adaptation:before}
}\qquad\qquad
\subfigure[Efficient overlay topology after adaptation averaging $\simeq 11$ delay units.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
  \label{figure:topology-adaptation:after}
}
\caption{A simplified example on how overlay topology adaptation may improve matters.}
\label{figure:topology-adaptation}
\end{figure}
%%

>>>>

While surveying efforts to overcome the mismatch problem
in the area of unstructured \p\ systems,
we came to identify four key methodologies 
utilized by the discussed approaches; they are:
\begin{enumerate}[\itshape i\upshape)]
  \item topology adaptation
  \item forwarding optimization
  \item caching and replication, and
  \item landmarking.
\end{enumerate}
%
A key characteristic shared by many of the presented techniques 
is that they do not strictly adhere to a single methodology.
As approaches attempt to address the challenging problem 
of topology mismatch, they resort to heuristics and so, 
they do not often furnish ``pure'' approaches 
(i.e., solutions that exclusively belong to 
one of the $4$ aforementioned methodologies). 
%%AD rephrased as above..
% An important aspect of many implementations, is that since the above
% methodologies use heuristics to solve topology mismatch, an
% intrinsically difficult problem, no single approach can offer a robust solution.
% Thus, many of the algorithms we surveyed are actually combinations of multiple
% such methodologies, in an attempt to refine the results of using just a single one.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For instance, \emph{mOverlay} combines topology adaptation 
with landmark binning 
while \emph{Gia} weaves together $3$ methodologies:  topology adaptation, 
forwarding optimization and caching. 
This combination inevitably leads to feature/performance trade-offs: for example, 
the effectiveness of a caching/replication component can be undermined
by a continuously adapting overlay that removes important links between peers.
If a technique is to be evaluated for a specific application domain,
its overall approach as well as its underlying used methodologies have to be considered.
%%AD rephrased as above... did not like much this "advised" - is not wrong BUT sounds funny
% For this reason, the reader is advised to critically approach either the overall
% algorithms or the separate methodologies when evaluating them for different
% application domains.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The cluster-based approaches, on the other hand, 
link physically-close nodes to each other. 
\emph{T2MC}, for example, uses trace-route logs and 
\emph{DDNO} exploits domain names to cluster nodes in proximity at peer join.
Further enhancements may include dynamic local restructuring of the overlay
graph through neighbor exchange like in \emph{PROP} or cycle-cut like in
\emph{DCMP} to achieve continuous adaptation throughout a peer's life-cycle.
Unfortunately, commonly used methods for proximity detection across Internet do
not always return
reliable results and therefore, mapping accuracy is not guaranteed. Specifically
for traceroute, its overhead is not negligible and routers or firewalls in a
network may have already been configured to disable traceroute response from
the start. 
The most problematic aspect of clustering, though, is
its nature per se. 
Limited connectivity among the various local domains can
significantly shrink the search scope, negatively affecting the query response
time that the \p\ user experiences. 
%%
Among others, \emph{DDNO} tries to 
balance the efficiency of the clustering approach 
with enhanced node connectivity by forcing half of each node's connections to be
with other, randomly selected, nodes.

As a methodology, topology adaptation, ultimately aims to reduce the 
average path traversing cost from one node to another.
In doing so, the methodology re-arranges the
overlay network so that it becomes a better fit 
for the underlying \emph{IP} network
as Figure~\ref{figure:topology-adaptation} depicts.
%%AD u always need a reference of a Figure in the text!!!
% as Figure~\ref{figure:topology-adaptation} depicts.
Despite the fact that the above is a plausible proposition,
it is not always advantageous though. For example, a topology adaptive
algorithm can exchange a slow edge in the overlay with a faster one
thus reducing the average latency of the network communication.
The pitfall here is that this new virtual link
can traverse a fast \emph{AS}--to--\emph{AS} link meaning that even though
message round-trip-time is reduced, it has actually additional cost in terms of
inter-\emph{ISP} communication accounting and management.
%%
\begin{figure}[ht]
\centering
\subfigure[Inefficient overlay topology averaging $\simeq 16$ delay units.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-before.pdf}
  \label{figure:topology-adaptation:before}
}\qquad\qquad
\subfigure[Efficient overlay topology after adaptation averaging $\simeq 11$ delay units.] {
  \includegraphics[scale=0.4]{img/pdf/topology-adaptation-after.pdf}
  \label{figure:topology-adaptation:after}
}
\caption{A simplified example on how overlay topology adaptation may improve matters.}
\label{figure:topology-adaptation}
\end{figure}
%%
